\chapter{Ethical considerations}
\label{chap:ethics}

Overall, there does not seem to be any major ethical concerns regarding the nature of 
this project. We have also investigated the licenses for the libaries we incorporated
within our tool, namely Mythril and the Solidity compiler, which respectively have the MIT and GPL-3.0
licenses, and should be permissive enough for our nature of work. 
However, there are still certain key points of consideration. First of all,
because the project deals with compiling Solidity code into 
EVM code, and generating the mappings between them, they may possibly reveal or
expose vulnerabilities within the compiler generated code. With Ethereum smart contracts 
nowadays holding up to millions of dollars worth of funds, if such information is revealed
to malicious actors, it could result in a permanent and irrevocable loss of funds.

In addition, this feature may also allow malicious actors to more easily obfuscate their code,
by possibly implanting fragments of malicious inline assembly into template contract codes,
claiming they serve another purpose. Then, users who do not analyse the code in detail and
simply reuse them may at best lead to bugs in execution, or at worst loss of funds.

Next, there is no guarantee provided by our project about the precision or reliability
of the gas analysis generated. Therefore, if the gas analysis is not accurate, and other projects
do not carefully consider the gas used by their contract calls and simply rely on estimates derived
from our tool, it may result in out of gas errors in future transactions, which may in turn
lead to funds being locked forever and lost. We therefore intend to waive any liabilities derived
from the use of our tool.
